# ---
# title: "EB - Markov chain Monte Carlo CalibMouseion Diagnosis"
# author: "Yu-Sheng Lin and Nan-Hung Hsieh" 
# date: "2023/02/05 (update: `r Sys.Date()`)"
# output: 
#  html_document:
# fig_caption: yes  
# ---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('..')
wd <- getwd()
knitr::opts_knit$set(root.dir =  wd)
```

# Prerequisites
```{r, include=FALSE}
source("MCSim/function.R")
getwd()
```

# Import the R packages
```{R message=FALSE, warning=FALSE}
library(tidyverse)
library(rstan)
library(bayesplot)
library(coda)
library(corrplot)
library(sensitivity)
library(pksensi)
library(data.table)
library(ggplot2)
library(GGally)
theme_set(theme_bw())
```

#### MCMC diagnostic analysis ###
```{r, eval=F}
Mousejob1 <- fread("outputs/EBMouseMCMC_4880.out")|> as.data.frame()
Mousejob2 <- fread("outputs/EBMouseMCMC_3365.out")|> as.data.frame()
Mousejob3 <- fread("outputs/EBMouseMCMC_5916.out")|> as.data.frame()
Mousejob4 <- fread("outputs/EBMouseMCMC_6734.out")|> as.data.frame()
Mouse_x <- mcmc_array(data = list(Mousejob1, Mousejob2, Mousejob3, Mousejob4))
```

# Size of output summary
```{r}
dim(Mouse_x)
```



# Andrew Gelman Scale Reduction Factor
```{r}
Mousemonitor <-monitor(Mouse_x, digits=4, probs = c(0.025, 0.5, 0.975), warmup = floor(dim(Mouse_x)[1]/2))
print(Mousemonitor)
```

# Dalaijamts MPSRF
```{r}
MousejobMPSRF1 = as.mcmc(fread("outputs/EBMouseMCMC_4880.out") |> as.data.frame(), header=T) 
MousejobMPSRF2 = as.mcmc(fread("outputs/EBMouseMCMC_3365.out") |> as.data.frame(), header=T) 
MousejobMPSRF3 = as.mcmc(fread("outputs/EBMouseMCMC_5916.out") |> as.data.frame(), header=T) 
MousejobMPSRF4 = as.mcmc(fread("outputs/EBMouseMCMC_6734.out") |> as.data.frame(), header=T) 

x1Mouse <- MousejobMPSRF1[,2:97 , drop=FALSE]
x2Mouse <- MousejobMPSRF2[,2:97 , drop=FALSE]
x3Mouse <- MousejobMPSRF3[,2:97 , drop=FALSE]
x4Mouse <- MousejobMPSRF4[,2:97 , drop=FALSE]

combined_Mouse = mcmc.list(x1Mouse,x2Mouse,x3Mouse,x4Mouse) 
gelman.diag(combined_Mouse, autoburnin=TRUE, multivariate=TRUE)
```

# traceplot and density diagnosis (every 10th sample from last half of MCMMC runs shown) #############
```{r, eval=F}
Mousejob1 <- fread("outputs/EBMouseMCMC_4880.out")|> as.data.frame()
Mousejob2 <- fread("outputs/EBMouseMCMC_3365.out")|> as.data.frame()
Mousejob3 <- fread("outputs/EBMouseMCMC_5916.out")|> as.data.frame()
Mousejob4 <- fread("outputs/EBMouseMCMC_6734.out")|> as.data.frame()
Mouse_x <- mcmc_array(data = list(Mousejob1, Mousejob2, Mousejob3, Mousejob4))
str <- ceiling(nrow(Mouse_x)/2) + 1
end <- nrow(Mouse_x)
j <- c(str:end) # discard burn-in
sample_x <- seq(100001, 200000, 10)
Mouse_x2 <- Mouse_x[sample_x,,]
pars_Pop <- c("M_lnPBC(1)",	"M_lnPFatC(1)",	"M_lnPLivC(1)",	"M_lnPRptC(1)", "M_lnPSptC(1)", "M_lnPLungC(1)",	"M_lnVmaxC(1)",	"M_lnKMC(1)",	"M_lnVmax2C(1)",	"M_lnKM2C(1)",	"M_lnVmaxLungC(1)",	"M_lnKMLungC(1)",	"M_lnVmaxRptC(1)",	"M_lnKMRptC(1)")

dp_Mouse <- bayesplot::mcmc_dens_overlay(Mouse_x2 [,,],  pars = pars_Pop,  facet_args = list(ncol= 3, nrow = 5))
tp_Mouse <-bayesplot::mcmc_trace(Mouse_x2 [,,], pars = pars_Pop,  facet_args = list(ncol= 3, nrow = 5))
ggplot2::ggsave(filename = "plots/suppl/Figure_S1_traceplot_Mouse.eps", 
                plot = tp_Mouse, 
                device = cairo_ps, 
                dpi = 1200, 
                width = 15,
                height = 10, 
                units = "cm") 
ggplot2::ggsave(filename = "plots/suppl/Figure_S4_density_Mouse.eps", 
                plot = dp_Mouse, 
                device = cairo_ps, 
                dpi = 1200, 
                width = 15,
                height = 10, 
                units = "cm")
```

# Correlation matrix of population posterior EB-specific parameters 
```{r}
sim1.1 <- fread("outputs/EBMouseMCMC_3365.out") 
sim2.1 <- fread("outputs/EBMouseMCMC_4880.out") 
sim3.1 <- fread("outputs/EBMouseMCMC_5916.out") 
sim4.1 <- fread("outputs/EBMouseMCMC_6734.out") 

SIM1 <- do.call(rbind, list(sim1.1))
SIM2 <- do.call(rbind, list(sim2.1))
SIM3 <- do.call(rbind, list(sim3.1))
SIM4 <- do.call(rbind, list(sim4.1))

Mouse_corr <-mcmc_array(list(SIM1, SIM2, SIM3, SIM4))
#dim(Mouse_corr)

pars_name <- dimnames(Mouse_corr)[[3]]
str <- which(pars_name == "M_lnVmaxC(1)")
end <- which(pars_name == "M_lnKMRptC(1)")
M_pars <- pars_name[str:end]
sample_x <- seq(100001, 200000, 10)
Parameter <- c(M_pars) %>%
  strsplit(split =  "\\(1\\)") %>% unlist()
# Correlation matrix
sum_chains <- length(sample_x)*4

Mouse_corr[sample_x,,M_pars] %>% matrix(nrow = sum_chains) %>%
  `colnames<-`(Parameter[1:8]) %>%
  ggcorr(nbreaks = 5, label = T, label_size = 4, label_round = 2, size = 4, label_alpha = TRUE) +
  theme_minimal() + labs(title = "Mouse") + theme(plot.title = element_text(size=26, hjust = 0.5)) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), title = element_blank()) +   theme(plot.margin=grid::unit(c(0,0,0,0), "mm")) 
ggsave("plots/suppl/Figure_S7_Correlation_matrix_Mouse.pdf", dpi = 1200, height = 10, width = 15, units="in")
```

```{r, include=FALSE}
clear()
```
