# ---
# title: "EB - Markov chain Monte Carlo CalibRation Diagnosis"
# author: "Yu-Sheng Lin et al" 
# date: "2023/02/05 (update: `r Sys.Date()`)"
# output: 
#  html_document:
# fig_caption: yes  
# ---
# setting up file path
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('..')
wd <- getwd()
knitr::opts_knit$set(root.dir =  wd)
```

# Prerequisites
```{r, include=FALSE}
source("MCSim/function.R")
getwd()
```

# Import the R packages
```{R message=FALSE, warning=FALSE}
library(tidyverse, quietly=T)
library(rstan, quietly=T)
library(bayesplot, quietly=T)
library(coda, quietly=T)
library(corrplot, quietly=T)
library(sensitivity, quietly=T)
library(pksensi, quietly=T)
library(data.table, quietly=T)
library(GGally, quietly=T)
library(ggplot2, quietly=T)
library(ggcorrplot, quietly=T)
theme_set(theme_bw())
```

#### MCMC diagnostic analysis ###
```{r, eval=F}
Humanjob1 <- fread("outputs/EBHumanMCMC_4880.out")|> as.data.frame()
Humanjob2 <- fread("outputs/EBHumanMCMC_3365.out")|> as.data.frame()
Humanjob3 <- fread("outputs/EBHumanMCMC_5916.out")|> as.data.frame()
Humanjob4 <- fread("outputs/EBHumanMCMC_6734.out")|> as.data.frame()
Human_x <- mcmc_array(data = list(Humanjob1, Humanjob2, Humanjob3, Humanjob4))
```

# Andrew Gelman Scale Reduction Factor
# ## 
# This section will produce results for posterior distributions for Supplemental Table S7 and S10 (output results from MCMC analysis)
# ## 
# From Supplemenal Table S7, one can use the equation in Supplemental Table S1 to calculate model parameters. Additional details are given in Supplemental Section 2.2 
# ## 
# For example,  PB (blood/air paritition coefficient) = PB0 × exp(lnPBC)  ----- this is can be seen in Supplemental Table S1 
# where PB0 is the species-specific baseline point estimate (28 for human) and lnPBC is the scaling (sampled) parameter of PBC in the log scale. 
# ##  
# PBC = exp(lnPBC), 
# Notethat at baseline (prior distribution, Table S4), lnPBC is assigned the baseline value of zero (0) prior to MCMC analysis (denoted as M_QLivC in 
# the code and output).  
# ##
# Thus, baseline (prior distribution),                              PB (prior blood/air paritition coefficient) = 28 × exp(0) = 28 for male mice   
# Upon completion of MCMC analysis (Supplemental Table S7),         PB (posterior blood/air paritition coefficient) = 28 × exp(-0.163) = 23.8 for male mice
# ##  
# For comparison, the prior vs posterior parameter distributions are shown in Table 3 (general physiological parameters) and Table 4 (EB-specific parameters)  
# ##  
# For Supplemental Table S10, residual error geometric standard deviation was calculated as exponential form of the square root of the error term 
# (e.g., “Ve_CV” for EB in blood) 
```{r}
Humanmonitor <-monitor(Human_x, digits=4, probs = c(0.025, 0.5, 0.975), warmup = floor(dim(Human_x)[1]/2))
print(Humanmonitor)
```

# Dalaijamts MPSRF
# This section will produce MPSRF values for Supplemental Figure S3 traceplot
```{r}
HumanjobMPSRF1 = as.mcmc(fread("outputs/EBHumanMCMC_4880.out") |> as.data.frame(), header=T) 
HumanjobMPSRF2 = as.mcmc(fread("outputs/EBHumanMCMC_3365.out") |> as.data.frame(), header=T) 
HumanjobMPSRF3 = as.mcmc(fread("outputs/EBHumanMCMC_5916.out") |> as.data.frame(), header=T) 
HumanjobMPSRF4 = as.mcmc(fread("outputs/EBHumanMCMC_6734.out") |> as.data.frame(), header=T) 

x1Human <- HumanjobMPSRF1[,2:126 , drop=FALSE]
x2Human <- HumanjobMPSRF2[,2:126 , drop=FALSE]
x3Human <- HumanjobMPSRF3[,2:126 , drop=FALSE]
x4Human <- HumanjobMPSRF4[,2:126 , drop=FALSE]

combined_Human = mcmc.list(x1Human,x2Human,x3Human,x4Human) 
gelman.diag(combined_Human, autoburnin=TRUE, multivariate=TRUE)
```

# traceplot and density diagnosis (every 10th sample from last half of MCMMC runs shown) #############
# This section will produce Supplemental Figure S3 and S6 traceplot
```{r, eval=F}
Humanjob1 <- fread("outputs/EBHumanMCMC_4880.out")|> as.data.frame()
Humanjob2 <- fread("outputs/EBHumanMCMC_3365.out")|> as.data.frame()
Humanjob3 <- fread("outputs/EBHumanMCMC_5916.out")|> as.data.frame()
Humanjob4 <- fread("outputs/EBHumanMCMC_6734.out")|> as.data.frame()
Human_x <- mcmc_array(data = list(Humanjob1, Humanjob2, Humanjob3, Humanjob4))
str <- ceiling(nrow(Human_x)/2) + 1
end <- nrow(Human_x)
j <- c(str:end) # discard burn-in
sample_x <- seq(100001, 200000, 10)
Human_x2 <- Human_x[sample_x,,]
pars_Pop <- c( "M_lnPBC(1)" , "M_lnPFatC(1)" , "M_lnPLivC(1)" , "M_lnPRptC(1)" , "M_lnPSptC(1)" , "M_lnPLungC(1)" , "M_lnVmaxC(1)" , "M_lnKMC(1)" , "M_lnVmax2C(1)" , "M_lnKM2C(1)",	"M_lnVmaxLungC(1)",	"M_lnKMLungC(1)",	"M_lnFmeteC(1)",	"M_lnKmaC(1)",	"M_lnKpgaC(1)")

dp_Human <- bayesplot::mcmc_dens_overlay(Human_x2 [,,],  pars = pars_Pop, facet_args = list(ncol= 3, nrow = 5))
tp_Human <-bayesplot::mcmc_trace(Human_x2 [,,], pars = pars_Pop,  facet_args = list(ncol= 3, nrow = 5))
ggplot2::ggsave(filename = "plots/suppl/Figure_S3_traceplot_Human.eps", 
                plot = tp_Human, 
                device = cairo_ps, 
                dpi = 1200, 
                width = 15,
                height = 10, 
                units = "cm")
ggplot2::ggsave(filename = "plots/suppl/Figure_S6_density_Human.eps", 
                plot = dp_Human, 
                device = cairo_ps, 
                dpi = 1200, 
                width = 15,
                height = 10, 
                units = "cm")
```

# Correlation matrix of population posterior EB-specific parameters 
# For Supplemental Figure S9
```{r}
sim1.1 <- fread("outputs/EBHumanMCMC_3365.out") 
sim2.1 <- fread("outputs/EBHumanMCMC_4880.out") 
sim3.1 <- fread("outputs/EBHumanMCMC_5916.out") 
sim4.1 <- fread("outputs/EBHumanMCMC_6734.out") 

SIM1 <- do.call(rbind, list(sim1.1))
SIM2 <- do.call(rbind, list(sim2.1))
SIM3 <- do.call(rbind, list(sim3.1))
SIM4 <- do.call(rbind, list(sim4.1))

Human_corr <-mcmc_array(list(SIM1, SIM2, SIM3, SIM4))
#dim(Human_corr)

pars_name <- dimnames(Human_corr)[[3]]
str <- which(pars_name == "M_lnVmaxC(1)")
end <- which(pars_name == "M_lnKpgaC(1)")
M_pars <- pars_name[str:end]
sample_x <- seq(100001, 200000, 10)
Parameter <- c(M_pars) %>%
  strsplit(split =  "\\(1\\)") %>% unlist()
# Correlation matrix
sum_chains <- length(sample_x)*4

Human_corr[sample_x,,M_pars] %>% matrix(nrow = sum_chains) %>%
  `colnames<-`(Parameter[1:9]) %>%
  ggcorr(nbreaks = 5, label = T, label_size = 4, label_round = 2, size = 4, label_alpha = TRUE) +
  theme_minimal() + labs(title = "Human") + theme(plot.title = element_text(size=26, hjust = 0.5)) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), title = element_blank()) +   theme(plot.margin=grid::unit(c(0,0,0,0), "mm")) 
ggsave("plots/suppl/Figure_S9_Correlation_matrix_Human.pdf", dpi = 1200, height = 10, width = 15, units="in")
```

```{r, include=FALSE}
clear()
```
